### 2022-05-03 code/analysis.Rmd
Convert to papaja + add supplement

---
title             : "A small male advantage in infant mental rotation"
shorttitle        : "Infant mental rotation"

author: 
  - name          : "Alexander Enge"
    affiliation   : "1,2"
    corresponding : "yes"
    address       : "Stephanstraße 1a, 04103 Leipzig"
    email         : "enge@cbs.mpg.de"
  - name          : "Shreya Kapoor"
    affiliation   : "1"
  - name          : "Anne-Sophie Kieslinger"
    affiliation   : "1"
  - name          : "Michael A. Skeide"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Research Group Learning in Early Childhood, Max Planck Institute for Human Cognitive and Brain Sciences"
  - id            : "2"
    institution   : "Department of Psychology, Humboldt-Universität zu Berlin"

authornote: |
  \addORCIDlink{Alexander Enge}{0000-0003-0100-2297}

  \addORCIDlink{Shreya Kapoor}{0000-0003-2619-8257}

  The data and code for this study are openly available at ???.
  We have no conflict of interest to disclose.
  There were no ethical concerns since we did not collect any new data.

abstract: |
  Mental rotation, the cognitive process of moving an object in mind to predict how it looks in a new orientation, is tightly coupled to intelligence, learning and educational achievement (Shepard & Metzler, 1971; Hegarty & Kozhevnikov, 1999; Johnshon & Bouchard, 2005).
  On average, males outperform females in mental rotation tasks (Linn & Petersen 1985, Voyer et al. 1995, Voyer 2011, Maeda & Yoon 2013).
  When such performance differences emerge during development, however, remains poorly understood (Lauer et al. 2019).
  Here we analyzed effect sizes derived from 59 experiments conducted in 1,798 infants aged 3 to 16 months.
  We robustly found that male infants looked longer at rotated objects than female infants.
  These findings indicate that the average male cognitive advantage in mental rotation already emerges in the first months of life.

keywords          : "mental rotation, spatial cognition, spatial ability, infants, children, gender, meta-analysis"
wordcount         : "X"

bibliography      : ["r-references.bib"]
csl               : "nature.csl"

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa7"
classoption       : "man,donotrepeattitle"
fontsize          : "10pt"
output            : papaja::apa6_pdf
header-includes:
  - \geometry{a4paper}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \renewcommand{\cellset}{\renewcommand{\arraystretch}{0.7}}
  - \captionsetup{font={stretch=1}, belowskip=15pt}
---

```{r, setup, include=FALSE}
# Load packages
library(papaja)
library(here)
library(ggridges)
library(tidyverse)
library(magrittr)
library(metafor)
library(brms)
library(bayestestR)
library(tidybayes)
library(cowplot)

# Global chunk options
knitr::opts_chunk$set(
  include = FALSE,
  message = FALSE,
  fig.height = 10,
  fig.width = 7.5,
  out.width = "100%",
  warning = FALSE
)

# Directory paths
data_dir <- here("data")
figures_dir <- here("results/figures")
tables_dir <- here("results/tables")

# Download literature search spreadsheet from Google Drive
excel_url <- "https://docs.google.com/spreadsheets/d/1xnC0NlQYTXavXaeqQidhi7UKb1pCGAl9Ry3h3Uz3mi0"
excel_file <- here(data_dir, "literature_search.xlsx")
# googledrive::drive_auth(use_oob = TRUE)
# googledrive::drive_download(excel_url, excel_file, overwrite = TRUE)

# Read sheets
screening <- readxl::read_excel(excel_file, sheet = "screening")
included <- readxl::read_excel(excel_file, sheet = "included", na = "NA")

# Add columns to the table of included experiments
included %>%
  mutate(

    # Add unique experiment identifier
    experiment = str_c(year, article, group, sep = ", "),

    # Add difference between condition means
    mean_diff = case_when(
      !is.na(mean_diff) ~ mean_diff,
      TRUE ~ mean_novel - mean_familiar
    ),
    # Add d_z from paired t test of condition means (Rosenthal, 1991)
    d_z_t = t / sqrt(sample_size),
    # Add d_z from ANOVA F value via conversion to a t value
    d_z_f = sqrt(f) / sqrt(sample_size) * sign_f,
    # Add d_z from mean and standard deviation of the difference
    d_z_diff = mean_diff / sd_diff,
    # Add d_av from mean difference and standard devations
    # See http://dx.doi.org/10.20982/tqmp.14.4.p242
    sd_av = sqrt((sd_novel^2 + sd_familiar^2) / 2),
    d_av = mean_diff / sd_av,
    # Add d from one-sample t test of novelty preference scores
    d_nov_pref = (nov_pref - 0.5) / sd_nov_pref,

    # Choose one type of outcome variable for each experiment
    di = case_when(
      # 1. If d was reported directed
      !is.na(d) ~ d,
      # 2. If a paired sample t test was reported
      !is.na(d_z_t) ~ d_z_t,
      # 3. If ANOVA was reported
      !is.na(d_z_f) ~ d_z_f,
      # 4. If the difference between means and its SD were reported
      !is.na(d_z_diff) ~ d_z_diff,
      # 5. If the individual condition means and their SDs were reported
      !is.na(d_av) ~ d_av,
      # 6. If a novelty preference score and its SD were reported
      !is.na(d_nov_pref) ~ d_nov_pref
    ),
    # Keep track which type of outcome measure was chosen for each article
    di_type = case_when(
      !is.na(d) ~ "d",
      !is.na(d_z_t) ~ "d_z_t",
      !is.na(d_z_f) ~ "d_z_f",
      !is.na(d_z_diff) ~ "d_z_diff",
      !is.na(d_av) ~ "d_av",
      !is.na(d_nov_pref) ~ "d_nov_pref",
      TRUE ~ "none"
    ) %>%
      factor(levels = c(
        "d", "d_z_t", "d_z_f", "d_z_diff", "d_av", "d_nov_pref", "none"
      )),

    # Apply small sample correction using Hedges' exact method
    # See http://dx.doi.org/10.20982/tqmp.14.4.p242
    dfi = 2 * (sample_size - 1),
    ji = exp(lgamma(dfi / 2) - log(sqrt(dfi / 2)) - lgamma((dfi - 1) / 2)),
    gi = di * ji,

    # Recode gender as a categorical (factor) variable for meta-regression
    gender = case_when(
      female_percent == 1.0 ~ "Female",
      female_percent == 0.0 ~ "Male",
      TRUE ~ "Mixed"
    ) %>% factor(levels = c("Mixed", "Female", "Male")),

    # Recode mean sample age in years (centered) for meta-regression
    age = age_mean / 365.25,
    age_c = age - mean(age, na.rm = TRUE),
    age_sd = age_sd / 365.25,

    # Recode task type as a categorical (factor) variale for meta-regression
    task = factor(task, levels = c("Habituation", "VoE")),

    # Combine gender and task into one column for plotting
    gender_task = factor(
      str_c(gender, task, sep = ", "),
      levels = c(
        "Female, Habituation",
        "Male, Habituation",
        "Mixed, Habituation",
        "Mixed, VoE"
      )
    )
  ) %>%
  # Order rows by experiment ID
  arrange(experiment) -> dat_all

# Overview of the different effect sizes
dat_all %>%
  select(
    article,
    group,
    gender_split,
    gi,
    di,
    di_type,
    d,
    d_z_t,
    d_z_f,
    d_z_diff,
    d_av,
    d_nov_pref
  ) %>%
  print(n = Inf)

# Compute standard error of Cohen's d based on assumed correlation
# See Hedges' formula on p. 253 in http://dx.doi.org/10.20982/tqmp.14.4.p242
# This will need a sensitivity analysis regarding the values of `r_assumed`
r_assumed <- 0.5
dat_all %>%
  mutate(
    ni = sample_size,
    vi = (dfi / (dfi - 2)) * ((2 * (1 - r_assumed)) / ni) *
      (1 + gi^2 * (ni / (2 * (1 - r_assumed)))) - (gi^2 / ji^2),
    sei = sqrt(vi)
  ) %>%
  filter(!is.na(gi)) -> dat_all_r

# Show number of experiments per gender grouping strategy
table(dat_all_r$gender_split)

# Extract split or mixed experiments only
dat_split <- filter(dat_all_r, gender_split %in% c("split", "split_only"))
dat_mixed <- filter(dat_all_r, gender_split %in% c("mixed", "mixed_only"))

# Combine both strategies, preferring split experiments
dat_both_split <- filter(
  dat_all_r, gender_split %in% c("split", "mixed_only", "split_only")
)

# Combine both strategies, preferring mixed experiments
dat_both_mixed <- filter(
  dat_all_r, gender_split %in% c("mixed", "mixed_only", "split_only")
)

# We go ahead with the last solution as to use the maximum amount of information
# while not biasing the gender results
dat <- dat_both_mixed
```

# Introduction

The cognitive ability to move visual object representations in mind for recognition across different orientations, known as mental rotation, emerges in the first three months of life (Johnson & Moore 2020, Moore & Johnson 2020).
Mental rotation is a key component of intelligence and a powerful predictor of learning outcome and educational achievement (Shepard & Metzler, 1971; Hegarty & Kozhevnikov, 1999; Johnshon & Bouchard, 2005).

Previous meta-analyses revealed that males solve mental rotation tasks slightly faster than females on average (Linn & Petersen 1985, Voyer et al. 1995, Voyer 2011, Maeda & Yoon 2013).
Effect sizes of this difference, however, are heterogeneous and often only medium (mean weighted g = 0.37–0.73).
Interestingly, a recent meta-analysis in 3-17-year-old children and adolescents suggests even only a small-to-medium male advantage (mean weighted g = 0.39) (Lauer et al. 2019).
Whether gender differences in mental rotation performance already emerge during infancy remains unknown.

In the present study we meta-analyzed 59 effect sizes derived from looking times in mental rotation tasks conducted by 1,798 infants (47.5% female) aged 3 to 16 months (mean = 7 months 14 days).
All tasks were embedded either into habituation experiments (40) or violation of expectation experiments (19) (Fantz 1964; Baillargeon et al. 1985).
These experiments comprised real world stimuli (e.g., toy objects, Antrilli & Wang, 2016), three-dimensional digital stimuli (e.g., cube figures; Moore & Johnson, 2008), or two-dimensional digital stimuli (e.g., digits; Quinn & Liben, 2008).
In habituation experiments, infants repeatedly saw an object until their looking times declined before they were presented with a mirror image of the object.
Longer looking times at the mirror image were taken as evidence that an infant still recognized the familiar object after rotation through the new orientation.
In violation of expectation experiments, infants were habituated to an object that was revolving repeatedly through a certain angle.
Then, either the familiar object or an unseen object was shown while they were revolving through a certain angle.
Subsequently, the object disappeared behind an occluder. Finally, the occluder was removed and either the object or its mirror object were shown.
Larger differences in looking times for the familiar object versus the unseen mirror image (at the new angle) were taken as evidence that an infant still recognized the familiar object after rotation through the new angle.
In contrast, looking times for both objects were expected to be similar if an infant did not recognize the familiar object from the new angle.

Following previous work in older populations we hypothesized that looking times in mental rotation experiments are on average significantly longer in male compared to female infants.
Effect sizes were assumed to be small.

# Results

## Mental rotation performance

```{r, meta_analysis}
# Specify priors
prior_c <- c(
  set_prior("normal(0, 1)", class = "b"),
  set_prior("cauchy(0, 0.3)", class = "sd")
)

# Run prior predictive simulation
n_iter <- 20000
n_warmup <- 1000
n_chains <- 4
res_prior <- brm(
  gi | se(sei) ~ 0 + Intercept + (1 | article / experiment),
  data = dat,
  prior = prior_c,
  sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  chains = n_chains,
  iter = n_iter,
  warmup = n_warmup,
  cores = n_chains,
  control = list(adapt_delta = 0.99),
  file = here("results", "models", "res_prior")
)
summary(res_prior)
# plot(res_prior)

# Run Bayesian multilevel model
res_brm <- update(
  res_prior,
  sample_prior = FALSE,
  file = here("results", "models", "res_brm")
)
summary(res_brm)
# plot(res_brm)

# # Summarise fixed effects (incl. Bayes factors)
# describe_posterior(
#   res_brm,
#   centrality = "mean", ci = 0.95, ci_method = "ETI",
#   test = c("p_direction", "rope", "bayesfactor"),
#   rope_range = c(-0.1, 0.1)
# )

# Compute by-experiment variance
var_experiment <- as.matrix(
  res_brm,
  variable = "sd_article:experiment__Intercept"
)^2
mean_qi(var_experiment)

# Compute by-article variance
var_article <- as.matrix(
  res_brm,
  variable = "sd_article__Intercept"
)^2
mean_qi(var_article)

# Compute intra-class correlation
var_total <- (var_experiment + var_article)
icc_draws <- var_article / var_total
mean_qi(icc_draws)
```

For our first meta-analytic model effect sizes were quantified as the standardized mean difference in infants’ looking times between novel and familiar rotated objects. Based on this effect size index we tested if there was evidence that infants can perform mental rotation. Across studies, infants indeed looked longer at novel rotated objects than at familiar rotated objects, with a standardized mean difference of g = 0.26, 95% credible interval (CrI) [0.12, 0.41] (Fig. 1). The probability for this effect being greater than zero was 99.9%. The heterogeneity of effect sizes was $sigma$2Experiment = 0.03, 95% CrI [0.00, 0.12], at the experiment level and $sigma$2Article = 0.17, 95% CrI [0.09, 0.29] at the article level. Therefore, approximately 86.7% of the heterogeneity between effect sizes was attributable to differences between experiments and 13.3% was attributable to differences between articles.

```{r, fig1, include=TRUE, fig.height=14, fig.width=12, fig.cap="(ref:fig1-caption)"}
# Helper functions for printing confidence intervals
print_num <- function(x, digits = 2) {
  x <- format(round(x, digits), trim = TRUE, nsmall = digits)
  x[x == "NA"] <- NA
  return(x)
}
print_ci <- function(est, lb, ub) {
  paste0(print_num(est), " [", print_num(lb), ", ", print_num(ub), "]")
}

# Get posterior draws for each experiment
draws <- epred_draws(res_brm, dat, ndraws = NULL) %>%
  mutate(experiment_f = factor(experiment))

# Get posterior draws for the meta-analytic effect
meta_draws <- spread_draws(res_brm, b_Intercept, ndraws = NULL)
meta_effect_size <- summary(res_brm)$fixed

# Create forest plot
n_experiments <- length(unique(dat$experiment))
dat %>%
  # Make sure the plot will be ordered alphabetically by experiment IDs
  arrange(experiment) %>%
  mutate(
    experiment_f = fct_rev(fct_expand(factor(experiment), "model")),
    # Show article labels only for the first experiment per article
    article = if_else(article == lag(article, default = ""), "", article),
    # Compute frequentist confidence intervals
    ci_lb = gi - qnorm(0.05 / 2, lower.tail = FALSE) * sqrt(vi),
    ci_ub = gi + qnorm(0.05 / 2, lower.tail = FALSE) * sqrt(vi),
    ci_print = print_ci(gi, ci_lb, ci_ub)
  ) %>%
  # Prepare plotting canvas
  ggplot(aes(x = gi, y = experiment_f)) +
  geom_vline(xintercept = seq(-3, 3, 0.5), color = "grey90") +
  # Add article and group labels as text on the left
  geom_text(aes(x = -9.9, label = article), hjust = 0) +
  geom_text(aes(x = -7.1, label = group_long), hjust = 0) +
  # Add experiment-specific effect sizes and CIs as text on the right
  geom_text(aes(x = 3.7, label = print_num(gi)), hjust = 1) +
  geom_text(aes(x = 4.4, label = print_num(ci_lb)), hjust = 1) +
  geom_text(aes(x = 5.1, label = print_num(ci_ub)), hjust = 1) +
  # Add Bayesian highest posterior density intervals for each experiment
  stat_interval(
    aes(x = .epred),
    data = draws,
    alpha = .8,
    point_interval = "mean_qi",
    .width = c(0.5, 0.95),
  ) +
  scale_color_grey(
    start = 0.85, end = 0.65,
    labels = as_mapper(~ scales::percent(as.numeric(.x)))
  ) +
  # Add experiment-specific effect sizes and CIs as dots with error bars
  geom_linerange(aes(xmin = ci_lb, xmax = ci_ub), size = 0.35) +
  geom_point(aes(size = ni), shape = 22, fill = "white") + # Or (1 / vi)?
  # Add posterior distribution for the meta-analytic effect
  stat_halfeye(
    aes(x = b_Intercept, y = -1),
    meta_draws,
    point_interval = "mean_qi",
    .width = c(0.95)
  ) +
  annotate(
    "text",
    x = c(-9.9, 3.7, 4.4, 5.1),
    y = -0.5,
    label = c(
      "Three-level model",
      print_num(meta_effect_size$Estimate),
      print_num(meta_effect_size$`l-95% CI`),
      print_num(meta_effect_size$`u-95% CI`)
    ),
    hjust = c(0, 1, 1, 1),
    fontface = "bold"
  ) +
  # Add headers for each of the for columns
  annotate( # Just to make the grid lines disappear
    "rect",
    xmin = -Inf,
    xmax = Inf,
    ymin = n_experiments + 0.5,
    ymax = Inf,
    fill = "white"
  ) +
  annotate(
    "text",
    x = c(-9.9, -7.1, 0, 3.7, 4.4, 5.1),
    y = n_experiments + 1.5,
    label = c(
      "Article",
      "Experiment",
      "Effect size plot",
      "Effect size",
      "Lower",
      "Upper"
    ),
    hjust = c(0, 0, 0.5, 1, 1, 1),
    fontface = "bold"
  ) +
  # Styling
  coord_cartesian(
    ylim = c(-1.5, n_experiments + 2), expand = FALSE, clip = "off"
  ) +
  scale_x_continuous(breaks = seq(-3, 3, 0.5)) +
  annotate("segment", x = -3.3, xend = 3.3, y = -1.5, yend = -1.5) +
  labs(
    x = expression("Hedges'" ~ italic("g")),
    size = "Sample size",
    color = "CrI level"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(family = "Helvetica", color = "black"),
    axis.text.y = element_blank(),
    axis.ticks.x = element_line(colour = "black"),
    axis.title.x = element_text(hjust = 0.67, margin = margin(b = -15)),
    axis.title.y = element_blank(),
    panel.grid = element_blank(),
    text = element_text(family = "Helvetica")
  )

# Save the plot
ggsave(here(figures_dir, "forest.pdf"), width = 12, height = 14)
```

(ref:fig1-caption) Mental rotation performance.
A Bayesian three-level meta-analysis provided evidence for mental rotation ability in infants.
White squares indicate the effect sizes (Hedges’ g) for infants’ mental rotation ability in all individual experiments and black lines their 95% confidence intervals.
Gray bars indicate the 95% and 50% Bayesian credible interval (CrI) based on a Bayesian three-level random-effects model.
Note that these intervals got shrunk towards the meta-analytic effect size because of partial pooling, that is, the three-level model regularizing the impact of experiments with small sample sizes and/or unrealistically large effect sizes.
The last line shows the meta-analytic effect size from the three-level model (black dot) together with its 95% CrI and its posterior distribution (gray curve).

## Effects of gender, age, and task type

```{r, meta_regression}
# Set numerical contrasts for factor variables
contrasts(dat$gender) <- MASS::contr.sdif(3)
contrasts(dat$task) <- MASS::contr.sdif(2)

# Specify priors
prior_c <- c(
  set_prior("normal(0, 1)", class = "b", coef = "Intercept"),
  set_prior("normal(0, 0.5)", class = "b"),
  set_prior("cauchy(0, 0.3)", class = "sd")
)

# Run prior predictive simulation
res_prior_reg <- brm(
  gi | se(sei) ~ 0 + Intercept + gender * age_c + task +
    (1 | article / experiment),
  data = dat,
  prior = prior_c,
  sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  chains = n_chains,
  iter = n_iter,
  warmup = n_warmup,
  cores = n_chains,
  control = list(adapt_delta = 0.99),
  file = here("results", "models", "res_prior_reg")
)
summary(res_prior_reg)
# plot(res_prior_reg)

# Run Bayesian multilevel model
res_brm_reg <- update(
  res_prior_reg,
  sample_prior = FALSE,
  file = here("results", "models", "res_brm_reg")
)
summary(res_brm_reg)
# plot(res_brm_reg)

# Summarise fixed effects (incl. Bayes factors)
describe_posterior(
  res_brm_reg,
  centrality = "mean", ci = 0.95, ci_method = "ETI",
  test = c("p_direction", "rope", "bayesfactor"),
  rope_range = c(-0.1, 0.1)
)

# Compute by-experiment variance
var_experiment_reg <- as.matrix(
  res_brm_reg,
  variable = "sd_article:experiment__Intercept"
)^2
mean_qi(var_experiment_reg)

# Compute by-article variance
var_article_reg <- as.matrix(
  res_brm_reg,
  variable = "sd_article__Intercept"
)^2
mean_qi(var_article_reg)

# Compute intra-class correlation
var_total_reg <- var_experiment_reg + var_article_reg
icc_reg <- var_article_reg / var_total_reg
mean_qi(icc_reg)

# Compute reduction in heterogeneity compared to the original meta-analysis
1 - mean(var_total_reg / var_total)
```

We next performed a meta-regression analysis to explain the heterogeneity of effect sizes and to test if the age of the gender of the infants, their age, and/or the type of mental rotation task influenced mental rotation performance (Fig. 2).
Indeed, there was evidence that experiments with all-male samples obtained better mental rotation performance than experiments with all-female samples, b = 0.48, 95% CrI [0.06, 0.91].
The probability for this effect being greater than zero was 98.5%.
Experiments with mixed-gender samples performed only slightly better than experiments than all-female samples, b = 0.13, 95% CrI [-0.20, 0.52].
The probability for this effect being greater than zero was 75.4%.
There was no evidence that the mean age of the infants influenced mental rotation performance, with a change per year of b = 0.05, 95% [-0.58, 0.68].
Similarly, there was no evidence for an interaction between gender and age.
Finally, there was weak evidence that violation of expectation tasks yielded larger effects than habituation tasks, b = 0.26, 95% [-0.08, 0.58].
The probability for this effect being greater than zero was 93.5%.

```{r, fig2, include=TRUE, fig.height=5, fig.width=12, fig.cap="(ref:fig2-caption)"}
# Get draws from the posterior distribution
draws_reg <- epred_draws(
  res_brm_reg,
  dat,
  ndraws = NA,
  re_formula = NA
)

# Use colors as proposed in https://arxiv.org/abs/2107.02270
habi_colors <- c(
  "Female, Habituation" = "#5790fc",
  "Male, Habituation" = "#f89c20",
  "Mixed, Habituation" = "#e42536"
)
voe_colors <- c("Mixed, VoE" = "#964a8b")

# Regression plot
dat %>%
  ggplot(aes(x = age * 12, y = gi)) +
  geom_hline(yintercept = seq(-1, 2, 0.5), color = "grey90") +
  annotate(
    "rect",
    xmin = 5.5, xmax = Inf, ymin = 1.25, ymax = Inf,
    color = NA, fill = "white"
  ) +
  # Regression lines and HPDIs - Habituation tasks
  stat_lineribbon(
    aes(y = .epred, color = gender_task),
    data = draws_reg,
    alpha = .4,
    point_interval = "mean_qi",
    .width = c(0.5, 0.95)
  ) +
  # Individual study effect sizes - Habituation tasks
  geom_point(aes(color = gender_task, size = ni), shape = 0) +
  scale_color_manual(
    values = habi_colors,
    breaks = names(habi_colors),
    labels = c("Females", "Males", "Mixed gender"),
    na.value = NA
  ) +
  guides(color = guide_legend(
    title = "Habituation tasks",
    override.aes = list(fill = NA),
    order = 1
  )) +
  # Regression lines and HPDIs - VoE tasks
  ggnewscale::new_scale_color() +
  stat_lineribbon(
    aes(y = .epred, color = gender_task),
    data = draws_reg,
    fill = NA,
    alpha = .4,
    point_interval = "mean_qi",
    .width = c(0.5, 0.95)
  ) +
  # Individual study effect sizes - VoE tasks
  geom_point(aes(color = gender_task, size = ni), shape = 0) +
  scale_color_manual(
    values = voe_colors,
    breaks = names(voe_colors),
    labels = c("Mixed gender"),
    na.value = NA
  ) +
  guides(color = guide_legend(title = "VoE tasks", order = 2)) +
  # Styling
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(limits = c(2.9, 16.1), breaks = seq(3, 16, 1)) +
  scale_y_continuous(limits = c(-1.2, 2.2), breaks = seq(-1, 2, 0.5)) +
  scale_fill_grey(
    start = 0.85, end = 0.65,
    labels = as_mapper(~ scales::percent(as.numeric(.x)))
  ) +
  labs(
    x = "Age (months)",
    y = expression("Hedges'" ~ italic("g")),
    fill = "CrI level",
    size = "Sample size"
  ) +
  theme_classic() +
  theme(
    axis.text = element_text(family = "Helvetica", color = "black"),
    axis.ticks = element_line(color = "black"),
    legend.direction = "vertical",
    legend.key = element_blank(),
    legend.position = "none",
    panel.grid = element_blank(),
    text = element_text(family = "Helvetica", color = "black")
  ) -> plot_reg

# Extract the legend so that we can put it inside the plotting area
plot_reg_legend <- get_legend(plot_reg + theme(legend.position = "top"))

# Define new labels for the regression coefficients
coef_colnames <- c(
  "Intercept" = "b_Intercept",
  "Females - mixed" = "b_gender2M1",
  "Males - females" = "b_gender3M2",
  "Age (per year)" = "b_age_c",
  "Habituation - VoE" = "b_task2M1",
  "(Females - Mixed) × age" = "b_gender2M1:age_c",
  "(Males - Females) × age" = "b_gender3M2:age_c"
)

# Plot posterior distributions of the regression coefficients
draws_reg_coef <- tidy_draws(res_brm_reg)
draws_reg_coef %>%
  select(all_of(coef_colnames)) %>%
  gather() %>%
  mutate(coef = factor(key, levels = names(coef_colnames))) %>%
  ggplot(aes(x = value, y = fct_rev(coef))) +
  geom_vline(xintercept = seq(-1, 1, 0.5), color = "grey90") +
  stat_halfeye(point_interval = "mean_qi", .width = c(0.5, 0.95)) +
  coord_cartesian(xlim = c(-1.25, 1.25), clip = "off") +
  scale_x_continuous(breaks = seq(-1, 1, 0.5)) +
  labs(x = expression("Regression weight (" * Delta * italic("g") * ")")) +
  annotate(
    "text",
    label = "Fixed effects", x = -3.175, y = 7.86, hjust = 0,
    family = "Helvetica", fontface = "bold", color = "black"
  ) +
  theme_minimal() +
  theme(
    axis.line.x = element_line(colour = "black"),
    axis.text.x = element_text(family = "Helvetica", color = "black"),
    axis.text.y = element_text(
      family = "Helvetica", color = "black", size = 3.88 * .pt,
      hjust = 0, vjust = -0.1, margin = margin(l = 20)
    ),
    axis.ticks.x = element_line(color = "black"),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid = element_blank(),
    text = element_text(family = "Helvetica", color = "black")
  ) -> plot_coef

# Combine plots
plot_grid(
  plot_reg, plot_coef,
  nrow = 1, rel_widths = c(3, 2), labels = "auto",
  label_size = 11, label_fontfamily = "Helvetica",
  label_x = 0.008, label_y = 0.99
) +
  draw_plot(plot_reg_legend, x = 0.385, y = 0.865, hjust = 0.5, vjust = 0.5)

# Save plot
ggsave(here(figures_dir, "regression_coef.pdf"), width = 12, height = 5)
```

(ref:fig2-caption) Effects of gender, age, and task task.
a Open squares show the effect size (Hedges’ g) for infants’ mental rotation ability in each of 59 individual experiments, color-coded according to the type of habituation task and the gender of the infants (blue = habituation task, all-female sample, yellow = habituation task, all-male sample, red = habituation task, mixed-gender sample, purple = violation of expectation [VoE] task, mixed-gender sample).
Lines indicate the best-fit regression estimates according to a Bayesian three-level meta-regression model and gray ribbons indicate their corresponding 95% and 50% credible interval (CrI).
b Fixed effect estimates from the Bayesian three-level meta-regression model are shown as black dots together with their 95% (thin black lines) and 50% (thick black lines) CrI.
Gray curves show the posterior distribution for each effect.

# Supplementary information

```{r, tabs1, include=TRUE}
# Helper functions to re-format age in years and/or months
format_days_months <- function(days) {
  days_per_month <- 30.417
  months <- days %/% days_per_month
  days_left <- round(days - months * days_per_month)
  str <- paste0(as.character(months), "m ", as.character(days_left), "d")
  str[is.na(days)] <- NA
  return(str)
}

# Save within-analysis table of experiments
dat %>%
  mutate(
    age_mean = format_days_months(age_mean),
    age_sd = str_c(as.character(round(age_sd)), "d"),
  ) %>%
  transmute(
    Article = if_else(article == lag(article, default = ""), "", article),
    Experiment = group_long,
    `Sample size` = sample_size,
    `Females` = round(female_percent * sample_size),
    Age = str_c(age_mean, "±", age_sd, sep = " "),
    Task = task,
    `Stimulus type` = str_to_sentence(stimuli_presentation),
    `Stimulus dimensions` = stimuli_dimensions
  ) -> tabS1

# Save the table
write_csv(tabS1, file = here(tables_dir, "within_experiments.csv"), na = "n/a")

# Display the table
apa_table(
  tabS1,
  caption = "Experiments included in the main analysis", label = "S1",
  landscape = TRUE
)
```

```{r, figs2, include=TRUE, fig.height=6, fig.width=12, fig.cap="(ref:figs2-caption)"}
# Extract convergence statistics from results
conv_stats <- with(summary(res_brm), tibble(
  parameter = c(
    "b_Intercept", "sd_article__Intercept", "sd_article:experiment__Intercept"
  ),
  name_expr = c(
    "Hedges ~ italic(g)", "italic(sigma)[article]", "italic(sigma)[experiment]"
  ),
  r_hat = print_num(c(fixed$Rhat, map_dbl(random, ~ .$Rhat)), digits = 4),
  bulk_ess = round(c(fixed$Bulk_ESS, map_dbl(random, ~ .$Bulk_ESS))),
  tail_ess = round(c(fixed$Tail_ESS, map_dbl(random, ~ .$Tail_ESS))),
  r_hat_expr = str_c("italic(R)[hat] == ", r_hat),
  bulk_ess_expr = str_c("italic(N)[eff(bulk)] == ", bulk_ess),
  tail_ess_expr = str_c("italic(N)[eff(tail)] == ", tail_ess),
  y_pos = c(0.7, 0.75, 0.8)
))

# Create trace plot
n_draws <- n_iter - n_warmup
bayesplot::mcmc_trace(
  res_brm,
  pars = conv_stats$parameter, facet_args = list(ncol = 1)
) +
  geom_text(
    aes(y = 0.45, label = name_expr, color = NA),
    conv_stats,
    y = 0.45, x = -800, color = "black", parse = TRUE, angle = 90
  ) +
  geom_text(
    aes(y = y_pos, label = r_hat_expr, color = NA),
    conv_stats,
    x = 13000, hjust = 1, color = "black", parse = TRUE,
  ) +
  geom_text(
    aes(y = y_pos, label = bulk_ess_expr, color = NA),
    conv_stats,
    x = 16000, hjust = 1, color = "black", parse = TRUE,
  ) +
  geom_text(
    aes(y = y_pos, label = tail_ess_expr, color = NA),
    conv_stats,
    x = 19000, hjust = 1, color = "black", parse = TRUE,
  ) +
  labs(x = "Sample number", y = NULL, color = "MCMC\nchain") +
  coord_cartesian(expand = FALSE, clip = "off") +
  scale_x_continuous(limits = c(0, n_draws), breaks = seq(0, n_draws, 2000)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  scale_color_grey(start = 0.0, end = 0.9) +
  theme_classic() +
  theme(
    axis.line = element_line(colour = "black"),
    axis.text = element_text(family = "Helvetica", color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.margin = margin(t = 10, l = 25),
    strip.background = element_blank(),
    strip.text = element_blank(),
    text = element_text(family = "Helvetica", color = "black")
  )

# Save the plot
ggsave(here(figures_dir, "trace.pdf"), width = 12, height = 6)
```

(ref:figs2-caption) Convergence checks.

```{r, frequentist_results}
# Three-level model
res_ml <- rma.mv(
  gi, vi,
  random = ~ 1 | article / experiment,
  data = dat,
  slab = experiment
)
print(res_ml)
# forest(res_ml)

# Check share of variance at each level of the model
round(res_ml$sigma2[1] / sum(res_ml$sigma2), 3) # Between-article (ICC)
round(res_ml$sigma2[2] / sum(res_ml$sigma2), 3) # Within-article (between-exp.)

# # Profile likelihodd plots for checking that REML has converged
# profile(res_ml)

# Meta-regression with gender, age, and task
res_reg <- rma.mv(
  gi, vi,
  mods = ~ gender * age_c + task,
  random = ~ 1 | article / experiment,
  data = dat,
  slab = experiment
)
print(res_reg)
# forest(res_full)
```

```{r, sensitivity_analysis, eval=FALSE, include=TRUE}
# Compute empirical estimate of the correlation between dependent samples
dat %>%
  mutate(
    # Get effect size from paired samples t-test, one sample t-test, or ANOVA
    d_diff = case_when(
      !is.na(d) ~ d,
      !is.na(d_z_t) ~ d_z_t,
      !is.na(d_z_f) ~ d_z_f,
      !is.na(d_z_diff) ~ d_z_diff
    ),
    # Compute empirical correlation
    # Basesd on the SD of the difference and the SDs within the two conditions
    sd_diff = mean_diff / d_diff,
    ri = (sd_diff^2 - sd_novel^2 - sd_familiar^2) /
      (-2 * sd_novel * sd_familiar)
  ) %>%
  pull(ri) -> ri_empirical

# Summarise the estimated correlations
summary(ri_empirical)

# Re-run the meta-analysis for different values of r_assumed
seq(-0.9, 0.9, by = 0.3) %>%
  set_names(., str_c("r = ", print_num(.))) %>%
  map(function(ri_sens) {

    # Re-compute standard errors of the effect sizes
    mutate(
      dat,
      vi = (dfi / (dfi - 2)) * ((2 * (1 - ri_sens)) / ni) *
        (1 + gi^2 * (ni / (2 * (1 - ri_sens)))) - (gi^2 / ji^2),
      sei = sqrt(vi)
    ) -> dat_ri_sens

    # Re-run meta-analysis
    update(res_brm, newdata = dat_ri_sens, refresh = 0)
  }) -> res_ri_sens

# Prior sensitivity analysis
list(
  "Intercept U(-10,10)" = c(
    set_prior("uniform(-10, 10)", class = "Intercept"),
    set_prior("cauchy(0, 0.3)", class = "sd")
  ),
  "Intercept N(0,0.2)" = c(
    set_prior("normal(0, 0.2)", class = "Intercept"),
    set_prior("cauchy(0, 0.3)", class = "sd")
  ),
  "SD U(0,10)" = c(
    set_prior("normal(0, 1)", class = "Intercept"),
    set_prior("uniform(0, 10)", class = "sd")
  ),
  "SD Student-t(10,0,0.2)" = c(
    set_prior("normal(0, 1)", class = "Intercept"),
    set_prior("student_t(10, 0, 0.2)", class = "sd")
  )
) %>%
  # Re-run the meta-analysis with different priors
  map(
    function(prior_sens) update(res_brm, prior = prior_sens, refresh = 0)
  ) -> res_prior_sens

# Create a table summarizing all sensitivity analyses
map2_dfr(
  c(res_ri_sens, res_prior_sens),
  c(names(res_ri_sens), names(res_prior_sens)),
  function(res, name) {

    # Compute ICC
    var_experiment <- as.matrix(
      res,
      variable = "sd_article:experiment__Intercept"
    )^2
    var_article <- as.matrix(
      res,
      variable = "sd_article__Intercept"
    )^2
    var_total <- var_experiment + var_article
    icc <- var_article / var_total
    icc_row <- mean_qi(icc) %>%
      rename(Estimate = y, `l-95% CI` = ymin, `u-95% CI` = ymax)

    # Create a data frame row of summary statistics
    with(summary(res), bind_rows(fixed, bind_rows(random))) %>%
      bind_rows(icc_row) %>%
      transmute(
        parameter = c("Hedges' g", "SD_{article}", "SD_{experiment}", "ICC"),
        Estimate = print_ci(Estimate, `l-95% CI`, `u-95% CI`)
      ) %>%
      pivot_wider(names_from = parameter, values_from = Estimate) %>%
      bind_cols(Simulation = name, .)
  }
) -> tab_sens

# Save the table
write_csv(tab_sens, file = here(tables_dir, "sensitivity_analysis.csv"))
```