### 2023-02-13 index.qmd
Update index and README

# Welcome {.unnumbered}

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/SkeideLab/meta_rotation.git/HEAD?urlpath=rstudio)
[![R version 4.1.2](https://img.shields.io/badge/R%20version-4.1.2-blue)](https://www.r-project.org)
[![GitHub workflow](https://img.shields.io/github/actions/workflow/status/SkeideLab/meta_rotation/docker-publish.yml)](https://github.com/SkeideLab/meta_rotation/actions/workflows/docker-publish.yml)
[![License](https://img.shields.io/github/license/SkeideLab/meta_rotation)](https://github.com/SkeideLab/meta_semantics/blob/main/LICENSE)

This webpage contains the code, data, and results for the following paper:

> Enge, A., Kapoor, S., Kieslinger, A.-S., & Skeide, M. A. (forthcoming). A meta-analysis of mental rotation in the first years of life. *Developmental Science*.

BibTeX citation:

```bibtex
@article{enge2023,
  title = {A Meta-Analysis of Mental Rotation in the First Years of Life},
  author = {Enge, Alexander and Kapoor, Shreya and Kieslinger, Anne-Sophie and Skeide, Michael A.},
  year = {forthcoming},
  journal = {Developmental Science}
}
```

![SkeideLab at MPI CBS](https://i.imgur.com/iPjdsfc.png)